---
title: "Plotting Freesurfers aparc areas"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting Freesurfers aparc areas}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=F}
knitr::opts_chunk$set(eval=T)
```

## Introduction
Plotting aparc areas from Freesurfers segmentations using ggplot2 is a handy way of displaying results or grouping of aparc data.

## Basic use
The function `ggaparc()` comes with the `MOAS` package for LCBC in-house use, but it is not limited in it's use to LCBC data, it can
be used in many different ways.
As it is based in the ggplot2 format, it is recommended to get a little familiarized with with [plotting data with ggplot2](http://r-statistics.co/Complete-Ggplot2-Tutorial-Part1-With-R-Code.html). 

Out-of-the-box, ggaparc() works without supplying any extra information. 
It will create a base plot of the aparc brain segmentations.

```{r start}
library(MOAS)
library(tidyverse)
library(ggplot2)
ggaparc()
```

The plot is kept as simple as possible, with as little extra information to ggplot as possible. 
This is done for it to be easy to extend the plot with the many ggplot options available.
For instance, you might want to only see the brain and it's segments, without any grid or axis information.
You can add ggplot2's `theme_void` to achieve this.

```{r void}
ggaparc() + theme_void()
```

You can supply some extra options to ggaparc, to alter the plot as you wish.
Maybe you want each area to have a different colour. 
ggaparc has a built column called `area` wich you can use as fill to achieve this.
**Note:** unlike other ggplots, before using `aes` you must indicate `mapping=aes`.

```{r fill}
ggaparc(mapping=aes(fill=area)) +
  theme_void()
```

Maybe you also want the contour lines to be black rather than white, and also a little thicker?

```{r fill2}
ggaparc(colour="black", size=.7, mapping=aes(fill=area)) +
  theme_void()
```

You can also zoom in on a single area and plot it alone.

```{r plot.area}
ggaparc(plot.area="transversetemporal") +
  theme_void()
```

Or you can plot several of them, despite not making much sense.

```{r plot.area2}
ggaparc(plot.area=c("transversetemporal", "insula","precentral","spc")) +
  theme_void()
```

If you want to plot the entire brain, but hightlight specific areas with colour, you will need to provide the function with some data.
The data **must** include a column names `area`, with area names corresponding to that of the aparc segmentations.
Here we create some data with 4 rows, and an `area` and `p` column. We provide the `p` column as the column
that dictates the colour of the area.

```{r datasupp}
someData = list(area=c("transverse", "smn","loc","spc"), p=sample(seq(0,.5,.001), 4)) %>% 
  as.data.frame(stringsAsFactors=F)

ggaparc(data=someData, mapping=aes(fill=p))

```

We can also change how this looks, by various ggplot functions
```{r datasupp2}
someData = list(area=c("transverse", "smn","loc","spc"), p=sample(seq(0,.5,.001), 4)) %>% 
  as.data.frame(stringsAsFactors=F)

ggaparc(data=someData, mapping=aes(fill=p)) +
  theme_void() +
  scale_fill_gradient(low="firebrick",high="goldenrod") +
  labs(title="Aparc plots rule")

```


If your data has different statistics for different groups, and you want to plot them all, we can use ggplot's `facet_wrap` or `facet_grid` for that.
The data needs to be in long format for this to work, like ggplot likes it.

```{r datasupp3}
someData = list(area=rep(c("transverse", "smn","loc","spc"),2), 
                p=sample(seq(0,.5,.001), 8),
                AgeG = c(rep("Young",4), rep("Old",4))
                ) %>% 
  as.data.frame(stringsAsFactors=F)

ggaparc(data=someData, mapping=aes(fill=p)) +
  facet_wrap(~AgeG) +
  theme(legend.position = "bottom")

```

We can change how this looks quite a bit. Maybe you want it on a dark background, and the areas without any data to be transparent?

```{r datasupp4}

ggaparc(data=someData, na.fill = "transparent",mapping=aes(fill=p)) +
  facet_wrap(~AgeG) +
  theme_dark() +
  theme(legend.position = "bottom",
        axis.text = element_blank(),
        axis.title = element_blank()
        )

```
