---
title: "Using the MOAS-package for LCBC custom functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the MOAS-package for LCBC custom functions}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval=F)
```

## Introduction
The Mother of all Spreadsheets (MOAS) data is a large data.frame/spreadsheet combining all information from participants in most of LCBC's studies. The spreadsheet is over 2000 columns wide, and 4000 rows long, and growing constantly. Navigating this file is tedious at times, and requires a lot of programming skills to handle.

The MOAS R-package is created witht he intent to share common operations/functions people may use on the MOAS, in order to get the data into the shape they need. This vignette is intended to showcase som of the functions currently available in the package that people may wish to use.

## How to use the package

### Basic usage
After the package is installed, the package needs to be called in every script or R instance to make it's functions available.

```{r "library", eval=TRUE}
library(MOAS)
library(tidyverse) #Recommended

```

The package functions mostly requires the actual MOAS data to work, or data created by subsetting the MOAS. 
The MOAS may be found in the MOAS data-foler in the Lagringshotell: `LCBC/Projects/Cross_projects/MOAS/Data/MOAS.RData`

Each project contained in the MOAS also has it's own MOAS-derived files at the top level of their data folder, in this type of format:
`LCBC/Projects/11_MemP/Data/11_MemP_Imaging_Long.RData`
These types of files are compatible with all functions in the package.

#### Loading the MOAS or MOAS-derived files
To work with the file, you must first get it into R's memory. 
For loading `.RData` files, which is an R-native file type, the simplest way is to use the `load()` function.
Remember to use the path to the file from **your current working directory**. 
Because `.RData` files are saved with the object name they had when saved, `.RData` can have very different names than expected when loading them. 
To handle this, we assign the loading of the file to an object `nm`, which will capture the name of the object.
We will then use the `get()` function to grab the data with name `nm` and assign it to the object `data`, so we have the data in an object with the name of our chosing.

```{r "Load RData"}
# Loads in the data, and also stores the name of the data.frame in the object `nm`
nm = load("LCBC/Projects/Cross_projects/MOAS/Data/MOAS.RData")

# Get the data with the name `nm` and assign it to `data`
data = get(nm)

```

You now have MOAS-data in your environment stored in the object `data`.

### Function documentation
Like all R-functions, the MOAS package functions have documentation that may be accessed by typing a question mark in the console, and the function you want to know more about.

```{r "help"}
# Base R
?anova
?t.test

# MOAS functions
?widen
?fs_lmm
```

#### The `widen()` function {#widen} 
This function will create a wider data.frame of the data you have, given some specification. 
This is often necessary if you want to use the data in SPSS, or for running ANOVAs etc.
To view the manual page of this package, simply type:

`widen()` has two mandatory inputs, the data to widen, and the column name to widen by. 

There are four options to widen the data by:  
1. "Subject_Timepoint"  
1. "Project_Wave"  
1. "Site_Name"  
1. "Site_Number"  

To widen data, simply type:
```{r "widen examplel1"}
widened_data = widen(data, by="Site_Name")
```

Using `by="Subject_Timepoint"` or `by="Project_Wave"` the data must have only _one line per participant and timepoint_. For double and triple scanned, this is not the case. 
In order to widen by these two options, you need to either run `widen()` with `by="Site_Name"` or `by="Site_Number"` first, or run the [`site_keeper()`](#site_keeper) function.

```{r "widen examplel2"}
widened_data1 = widen(data, by="Site_Name")
widened_data2 = widen(widened_data1, by="Project_Wave")
```


#### The `site_keeper()` function {#site_keeper} 
The `site_keeper()` function is intended for use when you wish to reduce double and triple scanned data to a single line from one of the scanners. 
This is necessary for all analyses not using the site as a variable of interest. 
To run the function, you must choose which scanner you wish to keep data from.   

There are four options:  
1. "long" - finds which scanner there is most data from, and keeps those (default) 
1. "ousAvanto" - keeps Avanto data  
1. "ousSkyra" - keeps Skyra data  
1. "ourPrisma" - keeps Prisma data  

**Note**: the operation only affects double and triple scan timepoints.
All participants retain the same amount of time points, but number of rows per double/triple scans is reduced to one.

```{r "site_keeper example1"}
simple_data = site_keeper(data, "long")
simple_data = site_keeper(data, "ousAvanto")
simple_data = site_keeper(data, "ourPrisma")
```

### Pipe ( %>% ) compatibility {#pipe}
All MOAS functions are pipe compatible, and should thus easily be incorporated into tidyverse-type syntax.

```{r "pipeExample"}
simple_data = data %>% 
  filter(Project_Name %in% c("MemC","MemP")) %>% 
  na.col.rm() %>% # see section on Utility functions
  site_keeper("ousPrisma") %>% 
  widen("Project_Wave")
  
```

### Utility functions
Some functions in the MOAS package are what we call utility functions.
They do not exclusively work on MOAS data, they will work well on other data.frames or vectors.
These funtions perform operations to help clean up data or perform simple and useful operations.


#### The `na.col.rm()` function {#na.col.rm} 
This function locates any column in a data.frame that has no observations (only contains `NA` or `NaN`), and removes it from the data.frame.
This particularly convenient if you have subsetted the rows of data to specific observations, and want to quickly remove any columns that no longer are of consequence for the data you have. 

```{r "na.col.rm"}
data2 = data %>% filter(Project_Name %in% "NCP") %>% na.col.rm()
```

#### The `count_chars()` function {#count_chars} 
You might want to count the number of times a specific character appears in a string variable, `count_chars()` can help you do this.

```{r "count_chars", eval=F}
strings = c("This is a vector","containing two strings")
strings %>% count_chars()

```

#### The `factor_times()` function {#factor_times} 
If you have a vector of "HH:MM" time specification, it might be convenient to create a factor of what time of day this would be. 
`factor_times()` takes a character vector, and produces a new carachte vector with up to four levels:  
- Morning  
- Afternoon  
- Evening  
- Night  

```{r "factor_times", eval=T, warning=F}
times_hhmm = c("22:40","19:30","08:21","02:47","11:45","13:12")
factor_times(times_hhmm)
```

#### The `calc_hour()` function {#calc_hour} 
If you have a vector of "HH:MM" time specification, it might be convenient to create a factor of what time of day this would be. 
`calc_hour()` takes a character time vector, a numeric vector of decimal hours.

```{r "calc_hour", eval=T, warning=F}
times_hhmm = c("22:40","19:30","08:21","02:47","11:45","13:12")
calc_hour(times_hhmm)
```


